A Large-scale Study of Polymorphism in R

### 2019-03-29 paper/main.tex
cleaned up stuff section, added bit about what data we collect, dunno if needed

%
%

%
\subsection{Granularity of Type Information}

As we mentioned in Section~\ref{sec:rlang}, R has several notions of
``type'', ranging from the results of \code{typeof} to metadata on values
obtainable via the \code{attributes} function.  Individually, these paint an
incomplete picture of the type of a value: For example, recall that
\code{typeof(5) == typeof(c(5, 5)) == "double"}.

To be as specific as possible, our analysis collects additional information
to create a more specific type than the result of any of R's reflection
functions.  For instance, if the analysis encounters {\tt typeof(x) ==
  "double"}, it will look at the length of {\tt x} to determine if {\tt x}
is a scalar or a vector, generating the annotation {\tt scalar/double} or
{\tt vector/double} accordingly.  One notable imperfection in this example
is that unit-length vectors will appear as scalars, but if some function was
called with unit-length vectors {\it as well as} larger ones, both scalar
and vector types will appear, and this distinction can be later collapsed.
Another example is that our analysis builds parametric list types by
capturing the types of list elements.  We also distinguish NAs ``in the
wild'' with a {\tt raw\_NA} type.

>>>>

%
%
%
%
\subsection{Granularity of Type Information}

When constructing a type for a value, we begin with the {\tt typeof} the value.
This alone paints an incomplete picture of the value, though, especially when it comes to vectors (recall that \code{typeof(5) == typeof(c(5, 5)) == "double"}).
In some cases, the class or attributes of a value might enlighten us, but in many cases we can be a little more precise.

To be as specific as possible, our analysis collects additional information
to create a more specific type than the result of any of R's reflection
functions.  The details of the information we collect follows.

	\item if the analysis encounters \code{ typeof(x) == "double"}, it will look at the length of {\tt x} to determine if {\tt x} is a scalar or a vector, generating the annotation {\tt scalar/double} or {\tt vector/double} accordingly.
	This is also true for integer, complex, raw, logical, and character {\tt typeof}s (i.e., the primitive vectorized data types).
	A notable imperfection here is is that unit length vectors will appear as scalars, so later in our analysis we ensure that argument signatures containing both scalar/X and vector/X are collapsed down to contain only vector/X.

	\item if the analysis encounters \code{typeof(x) == "list"}, we collect type information on all list elements in order to ascribe a general type to the contents of the list.
	To avoid undue slowdowns, we only collect the \code{typeof} the contents, and ascribe type {\tt any} if two or more {\tt typeof}s are observed.

	\item if the analysis encounters \code{NA}, we ascribe a unique {\tt raw\_NA} type.
	In R, \code{NA} inhabits all types, but \code{typeof(NA) == "logical"}.
	The analysis deals with these {\tt raw\_NA}s at a later stage.

	\item if the analysis encounters a matrix (i.e. the value has matrix class and has a dims and optionally dimnames attribute), we ascribe the type {\tt matrix}.

	\item similarly, if it encounters a data frame (i.e., class {\tt data.frame} with appropriate attributes), we ascribe the type {\tt data.frame}.
